SIMD\slash SIMT-vectorized Sparse Chemical Kinetic Jacobian\slash Thermo-chemical Source Term Evaluation

### 2018-03-10 paper.tex
fix

As demonstration of the type of issue we're discussing, a minimum working example has been created~\cite{nvidia_mwe} that demonstrates a failure of NVIDIA's OpenCL runtime corresponding to the GPU driver version \num{375.26} on a Tesla K40m GPU; by simply changing to another runtime (e.g., Intel), with no other code changes, the correct result can be found.
This provides a particularly vexing problem for the programmer, as there is often little that can be done to resolve the issue; thankfully, in this case we were able to upgrade the driver version to resolve the problem.
Further, as noted throughout this work certain transformations or code-generation patterns can break OpenCL execution\slash vectorization; e.g., the failure of POCL to achieve vectorization, Intel OpenCL's failure to vectorize the finite-difference Jacobian, the lack of speedup for the finite-difference Jacobian on the NVIDIA runtime for the \gpuold/ GPU.
Indeed the vectorization attempted by the OpenCL runtimes (or reasons for incorrectly vectorized code) are not always obvious to the user.
It is possible, particularly for the Intel OpenCL \slash POCL runtimes, that better performance might be achieved using so called ``explicit'' vectorization, i.e., through use of inbuilt vector types such as the \texttt{double8}.
This will be investigated in future works.

Thankfully, \texttt{loo.py} allows relatively easy switching between output languages; the most significant code change is building the wrapper that initializes\slash transfers memory and calls the kernel.
Indeed, the OpenMP code-generation built into \texttt{pyJac} is currently only capable of parallel-execution, but extending this platform to shallow\slash deep-vectorizations (via \texttt{loo.py} and compiler pragmas) is a key priority going forwards.
In addition, CUDA~\cite{NVIDIA:2018} and Intel's open-source OpenCL alternative ISPC~\cite{pharr2012ispc} have been significantly more reliable in previous works~\cite{Niemeyer:2016aa,CurtisGPU:2017} and (in the case of ISPC) initial testing.
The current deep-vectorization formulation would be executable for a CUDA target---as CUDA implements atomic double-precision floating-point operations---further recommending its use.
Extending execution to these targets to make CPU\slash GPU more reliable vectorization is also a future goal.

>>>>

As demonstration of the type of issue we're discussing, a minimum working example has been created~\cite{nvidia_mwe} that demonstrates a failure of NVIDIA's OpenCL runtime corresponding to the GPU driver version \num{375.26} on a Tesla K40m GPU; by simply changing to another runtime (e.g., Intel), with no other code changes, the correct result can be found.
This provides a particularly vexing problem for the programmer, as there is often little that can be done to resolve the issue; thankfully, in this case we were able to upgrade the driver version to resolve the problem.
Further, as noted throughout this work certain transformations or code-generation patterns can break OpenCL execution\slash vectorization; e.g., the failure of POCL to achieve vectorization, Intel OpenCL's failure to vectorize the finite-difference Jacobian, the lack of speedup for the finite-difference Jacobian on the NVIDIA runtime for the \gpuold/ GPU\todo{this went away?}.
Indeed the vectorization attempted by the OpenCL runtimes (or reasons for incorrectly vectorized code) are not always obvious to the user.
It is possible, particularly for the Intel OpenCL \slash POCL runtimes, that better performance might be achieved using so called ``explicit'' vectorization, i.e., through use of built-in vector types such as the \texttt{double8}.
This will be investigated in future works.

Thankfully, \texttt{loo.py} allows relatively easy switching between output languages; the most significant code change is building the wrapper that initializes\slash transfers memory and calls the kernel.
Indeed, the OpenMP code-generation built into \texttt{pyJac} is currently only capable of parallel-execution, but extending this platform to shallow\slash deep-vectorizations (via \texttt{loo.py} and compiler pragmas) is a key priority going forwards.
In addition, CUDA~\cite{NVIDIA:2018} and Intel's open-source OpenCL alternative ISPC~\cite{pharr2012ispc} have been significantly more reliable in previous works~\cite{Niemeyer:2016aa,CurtisGPU:2017} and (in the case of ISPC) initial testing.
The current deep-vectorization formulation would be executable for both CUDA and ISPC targets, as these languages implement double-precision atomic operations, further recommending their use.
Extending execution to these targets to make CPU\slash GPU more reliable vectorization is a future goal.
